---
title: "Stats 20, 22W -- Homework 2"
author: "Joanne Qiu -- 405369627"
date: "`r format(Sys.Date(), '%D')`"
output:
  html_document:
    toc: yes            # creates table of contents
    toc_depth: 4        # toc will include headers <= ####
    toc_float: yes      # toc always on left of page
    code_folding: show  # allows hiding of code
---

## Question 1

```{r Question 1 Background}
mixed1 <- c(TRUE, FALSE, FALSE, TRUE, 4, 0, 3)
mixed1
mixed2 <- c(TRUE, FALSE, FALSE, TRUE, 4, 0, "3")
mixed2
mixed3 <- c(c(TRUE, FALSE, FALSE, TRUE, 4, 0), "3")
mixed3
```

### (a)
<!-- ENTER YOUR ANSWER BELOW -->
Combining logical, numeric, and character vectors, `mixed2` results in a character vector that includes every element as a character. In contrast, `mixed3` concatenates two vectors. The first one combining logical and numeric vectors results in a numeric vector, so the logical vectors turn into 1 or 0. Then, when concatenating, the previous numeric vector and the character vector are combined, so the final result is a character vector. Thus `mixed2` has "TRUE" and "FALSE" while `mixed3` has "1" and "0".

### (b)
<!-- ENTER YOUR ANSWER BELOW, IT CAN BE JUST THE COMMAND. NO NEED TO STORE AN OBJECT -->
```{r Question 1b}
c(TRUE, FALSE, c(FALSE, TRUE, 4, 0), "3")
```

## Question 2

### (a)
<!-- ENTER YOUR ANSWER BELOW -->
`as.numeric(mixed2)` produces `r as.numeric(mixed2)` while `as.numeric(mixed3)` produces `r as.numeric(mixed3)`. In `mixed2`, "TRUE" and "FALSE" are not numbers, so they are printed as NA values. In contrast, the values of `mixed3` are all numeric, so they are printed directly as numbers.

### (b)
<!-- ENTER YOUR ANSWER BELOW -->
`as.logical(mixed2)` produces `r as.logical(mixed2)` while `as.logical(mixed3)` produces `r as.logical(mixed3)`. In `mixed2`, logical characters "TRUE" and "FALSE" are casted to logical values, but the numeric characters can not be casted. In `mixed3`, since all characters are numbers, they can not be casted to logical values. 

### (c)
<!-- ENTER YOUR ANSWER BELOW, IT CAN BE JUST THE COMMAND. NO NEED TO STORE AN OBJECT -->
```{r Question 2c}
as.logical(as.numeric(mixed3))
```

## Question 3

### (a)
<!-- ENTER YOUR ANSWERS BELOW, THEY CAN BE JUST THE COMMANDs. NO NEED TO STORE OBJECTs -->
```{r Question 3a}
seq(1, 25, by = 1) # (i) the seq() function

1:25 # (ii) the colon : operator

seq_len(25) # (iii) the seq_len() function

seq_along(1:25) # (iv) the seq_along() function
```

### (b)
<!-- ENTER YOUR ANSWERS BELOW, THEY CAN BE JUST THE COMMANDs. NO NEED TO STORE OBJECTs -->
```{r Question 3b}
seq(8, 2, length = 13) # (i)

rep(c(-1, 0, 3, 5), 5) # (ii)

rep(seq(5, 1), 5) # (iii)

rep(seq(5, 1), rep(5, 5)) # (iv)

rep(seq(10, 6), seq(1, 5)) # (v)
```

### (c)
<!-- ENTER YOUR ANSWER BELOW -->
The sequence is a combination of five different sequences consisting of five consecutive decreasing numbers with different starting points. Therefore, we can not use `rep()` to repeat the same sequence generated by `seq()` five times.

<!-- USE THE CODE CHUNK BELOW IF YOU'D LIKE -->
```{r Question 3c}
rep(seq(10, 6), 5) - rep(seq(0, 4), rep(5, 5))
```


## Question 4
<!-- USE THE CODE CHUNK BELOW -->
```{r Question 4a}
k = seq(0, 100)
sin(k * pi)
```
<!-- ENTER YOUR EXPLANATION BELOW -->
The output doesn't match what I would expect mathematically. In math, pi is an irrational number, and the sine of any integer times of pi should be conceptually zero. Here, the results are not zero because the default value of `pi` in R is `r pi` instead of the true value 3.1415926535.... 

## Question 5

### (a)
<!-- CREATE THE FUNCTION IN THE CODE CHUNK BELOW -->
<!-- NAME THE FUNCTION AND ITS INPUTS EXACTLY AS STATED IN THE ASSIGNENT -->
```{r Question 5a}
my_skew <- function(x) {
  xbar = mean(x)
  numerator = sqrt(length(x)) * sum((x - xbar)^3)
  denominator = (sum((x - xbar)^2))^(3 / 2)
  numerator / denominator
}
```

### (b)
<!-- USE THE CODE CHUNK BELOW -->
```{r Question 5b}
running_times <- c(51, 40, 57, 34, 47, 50, 50, 56, 41, 38)
my_skew(running_times)
```
<!-- ENTER YOUR EXPLANATION BELOW -->
The skewness coefficient is `r my_skew(running_times)`, which is less than 0, so the distribution of Chris Traeger's running times is slightly left-skewed.

### (c)
<!-- CREATE THE FUNCTION IN THE CODE CHUNK BELOW -->
<!-- NAME THE FUNCTION AND ITS INPUTS EXACTLY AS STATED IN THE ASSIGNENT -->
```{r Question 5c}
my_skew_one <- function(x) {
  xbar = mean(x)
  numerator = sqrt(length(x)) * (sum(x^3) - 3 * xbar * sum(x^2) + 2 * length(x) * xbar^3)
  denominator = (sum(x^2) - length(x) * xbar^2)^(3 / 2)
  numerator / denominator
}

my_skew(running_times)
my_skew_one(running_times)
```
<!-- ENTER YOUR EXPLANATION BELOW -->
`my_skew_one(running_times)` produces `r my_skew_one(running_times)`, which is the same as `r my_skew(running_times)` produced by `my_skew(running_times)`.

### (d)
<!-- USE THE CODE CHUNK BELOW -->
```{r Question 5d}
running_times * 10^10
my_skew(running_times)
my_skew_one(running_times)
```
<!-- ENTER YOUR EXPLANATION BELOW -->
The results are the same as the answers in (b) and (c). Multiplying the vector by 10^10 means to multiply every element in the vector by 10^10, so the mean is also multiplied by 10^10. By calculation, for both functions, to multiply x by 10^10 will result in multiplying both the numerator and the denominator by (10^10)^3. As the (10^10)^3s cancel each other out, the skewness doesn't change.

### (e)
<!-- USE THE CODE CHUNK BELOW IF YOU'D LIKE -->
```{r Question 5e}
my_skew(running_times) - my_skew_one(running_times)
```

<!-- ENTER YOUR ANSWER BELOW -->
Although the two functions produce very similar results, they are not identically the same. The two functions process the operations in different orders, so there are different rounding errors. Since the rounding errors are carried through calculation, the final results would be slightly different.

## Question 6

### (a)
<!-- CREATE THE FUNCTION IN THE CODE CHUNK BELOW -->
<!-- NAME THE FUNCTION AND ITS INPUTS EXACTLY AS STATED IN THE ASSIGNENT -->
```{r Question 6a}
my_length <- function(x) {
  index = seq_along(x) 
  index_1 = index / index # update every element in the list to 1
  sum(index_1)
}
```

### (b)
<!-- USE THE CODE CHUNK BELOW -->
```{r Question 6b}
my_length(numeric(0))
length(numeric(0))

my_length(seq(1, 99, by = 2))
length(seq(1, 99, by = 2))

my_length(c("friends", "waffles", "work"))
length(c("friends", "waffles", "work"))
```

## Question 7

### (a)
<!-- CREATE THE FUNCTION IN THE CODE CHUNK BELOW -->
<!-- NAME THE FUNCTION AND ITS INPUTS EXACTLY AS STATED IN THE ASSIGNENT -->
```{r Question 7a}
my_rev <- function(x) {
  x[length(x):0]
}
```

### (b)
<!-- USE THE CODE CHUNK BELOW -->
```{r Question 7b}
my_rev(numeric(0))
rev(numeric(0))

my_rev(seq(1, 99, by = 2))
rev(seq(1, 99, by = 2))

my_rev(c("friends", "waffles", "work"))
rev(c("friends", "waffles", "work"))
```

## Question 8

### (a)
<!-- CREATE THE FUNCTION IN THE CODE CHUNK BELOW -->
<!-- NAME THE FUNCTION AND ITS INPUTS EXACTLY AS STATED IN THE ASSIGNENT -->
```{r Question 8a}
front <- function(x, n) {
  x[0:(min(length(x), n))]
}

front(numeric(0), 6)
front(seq(1, 99, by = 2), 6)
front(c("friends", "waffles", "work"), 6)
front(c("friends", "waffles", "work"), 0)
```

### (b)
<!-- CREATE THE FUNCTION IN THE CODE CHUNK BELOW -->
<!-- NAME THE FUNCTION AND ITS INPUTS EXACTLY AS STATED IN THE ASSIGNENT -->
```{r Question 8b}
back <- function(x, n) {
  rev_x <- my_rev(x)
  rev_y <- front(rev_x, n)
  my_rev(rev_y)
}

back(numeric(0), 6)
back(seq(1, 99, by = 2), 6)
back(c("friends", "waffles", "work"), 6)
back(seq(1, 99, by = 2), 0)
```

### (c)
<!-- CREATE THE FUNCTION IN THE CODE CHUNK BELOW -->
<!-- NAME THE FUNCTION AND ITS INPUTS EXACTLY AS STATED IN THE ASSIGNENT -->
```{r Question 8c}
insert_at <- function(x, values, at) {
  beginning <- front(x, at - 1)
  ending <- back(x, max(length(x) - at + 1, 0))
  new_vec <- c(beginning, values, ending)
  new_vec
}
```

### (d)
<!-- USE THE CODE CHUNK BELOW -->
```{r Question 8d}
insert_at(1:5, c(0, 0, 0), at = 1)
insert_at(1:5, c(0, 0, 0), at = 3)
insert_at(1:5, c(0, 0, 0), at = 5)
insert_at(1:5, c(0, 0, 0), at = 6)
```
